A generalized quadratic estimate for random field nonstationarity \thanksref{code}

### 2016-02-28 paper/paper.tex
fresh figures, added a section on pull requests to README

\begin{figure}[H]
\includegraphics[height=6cm]{figure1.pdf}
\caption{
 These plots are intended to give an illustration  of the bias attenuation effect of the local invariant covariance structure (the specifics of the models are given in Section \ref{subsubsection: illustration of the bias reduction}).  The plot \textbf{at left} corresponds to the local invariant model and show estimation variance $\ell^2 C_{\ell}^{\text{var }\hat \phi}$ (\textbf{solid-green}), bias $\ell^2 C_{\ell}^{\text{bias }\hat \phi}$ (\textbf{dashed-red}) and the signal $\ell^2 C_{\ell}^{\phi\phi}$ (\textbf{dotted-black}). In contrast the plot \textbf{at right} shows their non local invariant counterparts.  Notice that the locally invariant second order bias is orders of magnitude smaller than the corresponding non local invariant bias, even though the functions $ C^{(1)}_k$ and $ C^{(2)}_k$ are similar to their non local invariant counterparts  $\widetilde C^{(1)}_k$ and $\widetilde C^{(2)}_k$. }
\label{Figure 1}
\end{figure}

@@

\begin{figure}
\includegraphics[height=16cm]{figure2.pdf}%
\caption{ This figure shows a simulation of a nonstationary phase random field $Z(x)$ (\textbf{top}), the quadratic estimate of  $\theta^\prime(x)$ (\textbf{middle}) and the spectral characterizations of estimation variance (\textbf{bottom-left}) and estimation bias (\textbf{bottom-right}). The details of the simulation are given in Section \ref{section: nonstat example d=1}. In the \textbf{middle plot} the \textbf{blue line} shows the value of $\theta^\prime(x)=\phi^{\prime\prime}(x)$ which characterizes the nonstationarity in $Z(x)$ through Definition \ref{def: nonstate phase}, the  \textbf{grey lines} show different quadratic estimates $\hat\theta^\prime(x)$ each one applied to an independent realizations of $Z(x)$ with the same $\theta(x)$ and the \textbf{dashed line} shows the empirical average of $\hat\theta^\prime(x)$ over 100 such realizations. In the \textbf{bottom-left plot}
The main purpose of this simulation is intended to illustrate the accuracy of the quadratic estimate and the ability of $C^{\text{var\,}\hat\phi}_{\bs \ell}$ and $C^{\text{bias\,}\hat\phi}_{\bs \ell}$ to approximate the empirical variance and bias of the estimate. A secondary goal of this figure is to also show that the fast approximation to $C^{\text{bias\,}\hat\phi}_{\bs \ell}$ is very accurate over a wide range of small wave numbers.}
\label{Figure 2 zx and var}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[height=16cm]{figure3.pdf}%
\caption{ An illustration of the quadratic estimate bias which results when $\theta^\prime(x)$ is large enough to exceed the \textbf{symmetric two-sided cut locus} discussed Section \ref{section: modeling xi, C and eta} (c.f. Claim \ref{claim: cut locus}). The \textbf{shaded region} shown in the \textbf{middle plot} corresponds to the interior of the two-sided cut locus. When the true  $\theta^\prime(x)$ (shown in \textbf{blue}) exits the symmetric two-sided cut locus, the quadratic estimate suffers from large bias, attenuating for negative $\theta^\prime(x)$ and amplifying for positive $\theta^\prime(x)$. Note that the only difference between this figure and Figure \ref{Figure 2 zx and var} is the parameter $t_0$ (c.f. Section \ref{section: modeling xi, C and eta}) which was reduced by a factor of $1/7$. This has the effect of shrinking the symmetric two-sided cut locus. All other parameters, including the random seed, are the same.}
\label{Figure 2 zx and var, extra bias}
\end{figure}

@@

@@

\begin{figure}[H]
\includegraphics[height=16cm]{figure4.pdf}%
\caption{An illustration of the quadratic estimate in $d=2$ for a nonstationary phase random field $Z(\bs x)$. The model for $Z(\bs x)$ is defined to have local variation in both the local smoothness of $Z(\bs x)$ and a local range parameter, where an increase of local smoothness corresponds reduction of local range and vice versa. The \textbf{top left image} shows the quadratic estimate $\hat\phi(\bs x)$ with the  ground truth $\phi(\bs x)$ shown in the \textbf{top right image} and the data shown in the \textbf{bottom right image}.  The \textbf{bottom left plot} shows the radial profile of  $|\bs \ell|^2C_{\bs \ell}^{\text{var }\hat\phi}$ (\textbf{solid green}), $|\bs \ell|^2C_{\bs \ell}^{\phi\phi}$ (\textbf{dotted black}) along with the fast approximation to $|\bs \ell|^2C_{\bs \ell}^{\text{bias }\hat\phi}$ (\textbf{dashed blue}) and the corresponding radially averaged empirical mean squared error per wavenumber (\textbf{dotted green}). See Section \ref{section: nonstat example d=2} for further simulation details.
 }
\label{Figure 3}
\end{figure}

@@

@@

\begin{claim}[{\bf Estimation variance}]\label{thm: quantify the var fluctuations of hat phi}
    Suppose $X(\bs x)$ is a mean-zero Gaussian random field with spectral density given by $C^{X\!X}_{\bs \ell}$.
    Then the spectral density $C_{\bs\ell}^{\text{\rm var }\hat{\phi}}$, defined by $E\big(\hat\phi_{\bs \ell}\{X,\!X\} \, \hat\phi_{\bs \ell^\prime}\{X,\!X\}^*\big) =  \delta^{\phantom{*}}_{\bs \ell-\bs \ell^\prime} C^{\text{var }\hat\phi}_{\bs \ell}$, satisfies
    \begin{align}
        C_{\bs\ell}^{\text{\rm var }\hat{\phi}}
        &= 2 A_{\bs\ell}^2
        \int
        \Bigl|\bs\xi_{\bs\ell}\!\cdot\!\bs C^{{(1)}}_{\bs k} - \bs\xi_{\bs\ell}\!\cdot\! \bs C^{{(1)}}_{\bs k+\bs \ell}\Bigr|^2
        \frac{C^{X\!X}_{\bs k+\bs\ell}}{(C^{ZZobs}_{\bs k+\bs\ell})^2}
        \frac{C^{X\!X}_{\bs k}}{(C^{ZZobs}_{\bs k})^2}
         \frac{d\bs k}{{(2\pi)}^{d}} \label{appendix, claim raw var} \\
        &= 2 A_{\bs\ell}^2  \sum_{p,q = 1}^d\bs\xi_{p,\bs\ell}\bs\xi^*_{q,\bs\ell}\int e^{-i\bs x\cdot\bs \ell} \left[2\mathscr A_{p,q}(\bs x)\mathscr B(\bs x)  - \mathscr C_{p}(\bs x)\mathscr C_{q}(\bs x) -  \mathscr D_{p}(\bs x)\mathscr D_{q}(\bs x)\right]  \frac{d\bs x}{{(2\pi)}^{d}} \label{appendix, claim fast var}
    \end{align}
    for all $\bs \ell\neq 0$
    where $\mathscr A_{p,q, \bs \ell}:= \bs C^{(1)}_{p,\bs \ell} \bs C^{{(1)}^*}_{q,\bs \ell} C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs \ell})^2$, $\mathscr B_\ell:=  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$, $\mathscr C_{p,\bs\ell}:= \bs C^{(1)}_{p,\bs\ell}  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$ and $\mathscr D_{p,\bs\ell}:= \bs C^{{(1)}^*}_{p,\bs\ell}  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$.
    Moreover, if $C_{\bs\ell}^{X\!X} = C^{ZZobs}_{\bs\ell}$  one obtains
    \begin{equation}
    C_{\bs\ell}^{\text{\rm var }\hat{\phi}} =  2{(2\pi)}^{-d/2} A_{\bs\ell}.
    \end{equation}
\end{claim}

\begin{proof}
    Recall Definition \ref{def: quad est applied to X} which states that $\hat\phi_{\bs \ell}\{X,\!X\}$ denotes the quadratic estimate applied to data $X(\bs x)$. Therefore 
    \begin{align}
        E\big(\hat\phi_{\bs \ell}\{X,\!X\} \, \hat\phi_{\bs \ell^\prime}\{X,\!X\}^*\big)
        & = A_{\bs \ell}A_{\bs \ell^\prime}
        \iint
        {\Big(\bs\xi_{\bs \ell} \!\cdot\!\bs C^{{(1)}}_{\bs k} - \bs\xi_{\bs \ell} \!\cdot\!\bs C^{{(1)}}_{\bs k+\bs \ell}\Big)}^{\! *}
        {\Big(\bs\xi_{\bs \ell^\prime} \!\cdot\!\bs C^{{(1)}}_{\bs k^\prime} - \bs\xi_{\bs \ell^\prime} \!\cdot\!\bs C^{{(1)}}_{\bs k^\prime+\bs \ell^\prime}\Big)}^{}\nonumber\\
        &\qquad\qquad\qquad
        \times
        \frac{E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell}X_{\bs k^\prime}\bigr)}{C^{ZZobs}_{\bs k+\bs \ell}C^{ZZobs}_{\bs k}C^{ZZobs}_{\bs k^\prime+\bs \ell^\prime}C^{ZZobs}_{\bs k^\prime}}
        \frac{d\bs kd\bs k^\prime}{{(2\pi)}^{d}}\label{wicks term in Cvar}
    \end{align}
    Now expanding the above fourth moment, using Wick's theorem (also called Isserlis's Theorem) \cite{wick1950evaluation, isserlis1916certain} and the Gaussianity of $X(\bs x)$, one obtains
    \begin{align*}
        E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell}X_{\bs k^\prime}\bigr)
        &=C^{X\!X}_{\bs k+\bs \ell}C^{X\!X}_{-\bs k^\prime}\bigl( \delta_{\bs\ell}\delta_{\bs\ell^\prime} +  \delta_{\bs k-\bs k^\prime + \bs \ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime} +  \delta_{\bs k + \bs k^\prime +\bs\ell}\delta_{\bs k + \bs k^\prime +\bs\ell}   \bigr) \\
        &=C^{X\!X}_{\bs k+\bs\ell}C^{X\!X}_{-\bs k^\prime}\bigl( \delta_{\bs\ell}\delta_{\bs\ell^\prime} +  \delta_{ 
        \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime} +  \delta_{\bs\ell-\bs\ell^\prime}\delta_{\bs k+\bs k^\prime +\bs\ell}   \bigr).
    \end{align*}
    The term $\delta_{\bs\ell}\delta_{\bs\ell^\prime}$ is only nonzero  when $\bs\ell=\bs\ell^\prime=0$. Furthermore, by a change of variables, one can see that effect of the terms $\delta_{ \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime}$ and $ \delta_{\bs\ell-\bs\ell^\prime}\delta_{\bs k+\bs k^\prime +\bs \ell}$ in (\ref{wicks term in Cvar}) are identical. Therefore assuming $\bs \ell \neq 0$ and replacing  $E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell}X_{\bs k^\prime}\bigr)$ in (\ref{wicks term in Cvar}) with $2C^{X\!X}_{\bs k+\bs\ell}C^{X\!X}_{-\bs k^\prime}\delta_{ \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime}$ gives (\ref{appendix, claim raw var}). A similar method to the derivation for (\ref{Aell derivation 11}) can be used to establish (\ref{appendix, claim fast var}). Finally, if one replaces $C^{X\!X}_{\bs k}$ in (\ref{appendix, claim raw var}) with the marginal spectral density $C^{ZZobs}_{\bs k}$ then using (\ref{Aell derivation 11}) one obtains $C_{\bs\ell}^{\text{\rm var }\hat{\phi}} =  2{(2\pi)}^{-d/2} A_{\bs\ell}$.

>>>>

\begin{figure}[H]
\includegraphics[height=6.2cm]{figure1.pdf}
\caption{
 These plots are intended to give an illustration  of the bias attenuation effect of the local invariant covariance structure (the specifics of the models are given in Section \ref{subsubsection: illustration of the bias reduction}).  The plot \textbf{at left} corresponds to the local invariant model and show estimation variance $\ell^2 C_{\ell}^{\text{var }\hat \phi}$ (\textbf{solid-green}), bias $\ell^2 C_{\ell}^{\text{bias }\hat \phi}$ (\textbf{dashed-red}) and the signal $\ell^2 C_{\ell}^{\phi\phi}$ (\textbf{dotted-black}). In contrast the plot \textbf{at right} shows their non local invariant counterparts.  Notice that the locally invariant second order bias is orders of magnitude smaller than the corresponding non local invariant bias, even though the functions $ C^{(1)}_k$ and $ C^{(2)}_k$ are similar to their non local invariant counterparts  $\widetilde C^{(1)}_k$ and $\widetilde C^{(2)}_k$. }
\label{Figure 1}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[height=15cm]{figure2.pdf}%
\caption{ This figure shows a simulation of a nonstationary phase random field $Z(x)$ (\textbf{top}), the quadratic estimate of  $\theta^\prime(x)$ (\textbf{middle}) and the spectral characterizations of estimation variance (\textbf{bottom-left}) and estimation bias (\textbf{bottom-right}). The details of the simulation are given in Section \ref{section: nonstat example d=1}. In the \textbf{middle plot} the \textbf{blue line} shows the value of $\theta^\prime(x)=\phi^{\prime\prime}(x)$ which characterizes the nonstationarity in $Z(x)$ through Definition \ref{def: nonstate phase}, the  \textbf{grey lines} show different quadratic estimates $\hat\theta^\prime(x)$ each one applied to an independent realizations of $Z(x)$ with the same $\theta(x)$ and the \textbf{dashed line} shows the empirical average of $\hat\theta^\prime(x)$ over 100 such realizations. In the \textbf{bottom-left plot}
The main purpose of this simulation is intended to illustrate the accuracy of the quadratic estimate and the ability of $C^{\text{var\,}\hat\phi}_{\bs \ell}$ and $C^{\text{bias\,}\hat\phi}_{\bs \ell}$ to approximate the empirical variance and bias of the estimate. A secondary goal of this figure is to also show that the fast approximation to $C^{\text{bias\,}\hat\phi}_{\bs \ell}$ is very accurate over a wide range of small wave numbers.}
\label{Figure 2 zx and var}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[height=15cm]{figure3.pdf}%
\caption{ An illustration of the quadratic estimate bias which results when $\theta^\prime(x)$ is large enough to exceed the \textbf{symmetric two-sided cut locus} discussed Section \ref{section: modeling xi, C and eta} (c.f. Claim \ref{claim: cut locus}). The \textbf{shaded region} shown in the \textbf{middle plot} corresponds to the interior of the two-sided cut locus. When the true  $\theta^\prime(x)$ (shown in \textbf{blue}) exits the symmetric two-sided cut locus, the quadratic estimate suffers from large bias, attenuating for negative $\theta^\prime(x)$ and amplifying for positive $\theta^\prime(x)$. Note that the only difference between this figure and Figure \ref{Figure 2 zx and var} is the parameter $t_0$ (c.f. Section \ref{section: modeling xi, C and eta}) which was reduced by a factor of $1/7$. This has the effect of shrinking the symmetric two-sided cut locus. All other parameters, including the random seed, are the same.}
\label{Figure 2 zx and var, extra bias}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[height=15cm]{figure4.pdf}%
\caption{An illustration of the quadratic estimate in $d=2$ for a nonstationary phase random field $Z(\bs x)$. The model for $Z(\bs x)$ is defined to have local variation in both the local smoothness of $Z(\bs x)$ and a local range parameter, where an increase of local smoothness corresponds reduction of local range and vice versa. The \textbf{top left image} shows the quadratic estimate $\hat\phi(\bs x)$ with the  ground truth $\phi(\bs x)$ shown in the \textbf{top right image} and the data shown in the \textbf{bottom right image}.  The \textbf{bottom left plot} shows the radial profile of  $|\bs \ell|^2C_{\bs \ell}^{\text{var }\hat\phi}$ (\textbf{solid green}), $|\bs \ell|^2C_{\bs \ell}^{\phi\phi}$ (\textbf{dotted black}) along with the fast approximation to $|\bs \ell|^2C_{\bs \ell}^{\text{bias }\hat\phi}$ (\textbf{dashed blue}) and the corresponding radially averaged empirical mean squared error per wavenumber (\textbf{dotted green}). See Section \ref{section: nonstat example d=2} for further simulation details.
 }
\label{Figure 3}
\end{figure}

@@

\begin{claim}[{\bf Estimation variance}]\label{thm: quantify the var fluctuations of hat phi}
    Suppose $X(\bs x)$ is a mean-zero Gaussian random field with spectral density given by $C^{X\!X}_{\bs \ell}$.
    Then the spectral density of $\hat\phi_{\bs \ell}\{X,\!X\}$ (c.f. Definition \ref{def: quad est applied to X}), which satisfies $\delta^{\phantom{*}}_{\bs \ell-\bs \ell^\prime}C^{\text{\rm var }\hat\phi}_{\bs \ell} = E\big(\hat\phi_{\bs \ell}\{X,\!X\} \, \hat\phi_{\bs \ell^\prime}\{X,\!X\}^*\big) $, is given as follows
    \begin{align}
        C_{\bs\ell}^{\text{\rm var }\hat{\phi}}
        &= 2 A_{\bs\ell}^2
        \int
        \Bigl|\bs\xi_{\bs\ell}\!\cdot\!\bs C^{{(1)}}_{\bs k} - \bs\xi_{\bs\ell}\!\cdot\! \bs C^{{(1)}}_{\bs k+\bs \ell}\Bigr|^2
        \frac{C^{X\!X}_{\bs k+\bs\ell}}{(C^{ZZobs}_{\bs k+\bs\ell})^2}
        \frac{C^{X\!X}_{\bs k}}{(C^{ZZobs}_{\bs k})^2}
         \frac{d\bs k}{{(2\pi)}^{d}} \label{appendix, claim raw var} \\
        &= 2 A_{\bs\ell}^2  \sum_{p,q = 1}^d\bs\xi_{p,\bs\ell}\bs\xi^*_{q,\bs\ell}\int e^{-i\bs x\cdot\bs \ell} \left[2\mathscr A_{p,q}(\bs x)\mathscr B(\bs x)  - \mathscr C_{p}(\bs x)\mathscr C_{q}(\bs x) -  \mathscr D_{p}(\bs x)\mathscr D_{q}(\bs x)\right]  \frac{d\bs x}{{(2\pi)}^{d}} \label{appendix, claim fast var}
    \end{align}
    for all $\bs \ell\neq 0$
    where $\mathscr A_{p,q, \bs \ell}:= \bs C^{(1)}_{p,\bs \ell} \bs C^{{(1)}^*}_{q,\bs \ell} C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs \ell})^2$, $\mathscr B_\ell:=  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$, $\mathscr C_{p,\bs\ell}:= \bs C^{(1)}_{p,\bs\ell}  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$ and $\mathscr D_{p,\bs\ell}:= \bs C^{{(1)}^*}_{p,\bs\ell}  C^{X\!X}_{\bs\ell} / (C^{ZZobs}_{\bs\ell})^2$.
    Moreover, if $C_{\bs\ell}^{X\!X} = C^{ZZobs}_{\bs\ell}$  one obtains
    \begin{equation}
    C_{\bs\ell}^{\text{\rm var }\hat{\phi}} =  2{(2\pi)}^{-d/2} A_{\bs\ell}.
    \end{equation}
\end{claim}

\begin{proof}
    Recall Definition \ref{def: quad est applied to X} which states that $\hat\phi_{\bs \ell}\{X,\!X\}$ denotes the quadratic estimate applied to data $X(\bs x)$. Therefore 
    \begin{align}
        E\big(\hat\phi_{\bs \ell}\{X,\!X\} \, \hat\phi_{\bs \ell^\prime}\{X,\!X\}^*\big)
        & = A_{\bs \ell}A_{\bs \ell^\prime}
        \iint
        {\Big(\bs\xi_{\bs \ell} \!\cdot\!\bs C^{{(1)}}_{\bs k} - \bs\xi_{\bs \ell} \!\cdot\!\bs C^{{(1)}}_{\bs k+\bs \ell}\Big)}^{\! *}
        {\Big(\bs\xi_{\bs \ell^\prime} \!\cdot\!\bs C^{{(1)}}_{\bs k^\prime} - \bs\xi_{\bs \ell^\prime} \!\cdot\!\bs C^{{(1)}}_{\bs k^\prime+\bs \ell^\prime}\Big)}^{}\nonumber\\
        &\qquad\qquad\qquad
        \times
        \frac{E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell^\prime}X_{\bs k^\prime}\bigr)}{C^{ZZobs}_{\bs k+\bs \ell}C^{ZZobs}_{\bs k}C^{ZZobs}_{\bs k^\prime+\bs \ell^\prime}C^{ZZobs}_{\bs k^\prime}}
        \frac{d\bs kd\bs k^\prime}{{(2\pi)}^{d}}.\label{wicks term in Cvar}
    \end{align}
    Now expanding the above fourth moment, using Wick's theorem (also called Isserlis's Theorem) \cite{wick1950evaluation, isserlis1916certain} and the Gaussianity of $X(\bs x)$, one obtains
    \begin{align*}
        E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell^\prime}X_{\bs k^\prime}\bigr)
        &=C^{X\!X}_{\bs k+\bs \ell}C^{X\!X}_{-\bs k^\prime}\bigl( \delta_{\bs\ell}\delta_{\bs\ell^\prime} +  \delta_{\bs k-\bs k^\prime + \bs \ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime} +  \delta_{\bs k + \bs k^\prime +\bs\ell}\delta_{\bs k + \bs k^\prime +\bs\ell^\prime}   \bigr) \\
        &=C^{X\!X}_{\bs k+\bs\ell}C^{X\!X}_{-\bs k^\prime}\bigl( \delta_{\bs\ell}\delta_{\bs\ell^\prime} +  \delta_{ 
        \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime} +  \delta_{\bs\ell-\bs\ell^\prime}\delta_{\bs k+\bs k^\prime +\bs\ell^\prime}   \bigr).
    \end{align*}
    The term $\delta_{\bs\ell}\delta_{\bs\ell^\prime}$ is only nonzero  when $\bs\ell=\bs\ell^\prime=0$. Furthermore, by a change of variables, one can see that effect of the terms $\delta_{ \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime}$ and $ \delta_{\bs\ell-\bs\ell^\prime}\delta_{\bs k+\bs k^\prime +\bs \ell^\prime}$ in (\ref{wicks term in Cvar}) are identical. Therefore assuming $\bs \ell \neq 0$ and replacing  $E\bigl(X_{\bs k+\bs \ell}X_{-\bs k}X_{-\bs k^\prime-\bs\ell^\prime}X_{\bs k^\prime}\bigr)$ in (\ref{wicks term in Cvar}) with $2C^{X\!X}_{\bs k+\bs\ell}C^{X\!X}_{-\bs k^\prime}\delta_{ \bs\ell-\bs\ell^\prime}\delta_{\bs k-\bs k^\prime}$ gives (\ref{appendix, claim raw var}). A similar method to the derivation for (\ref{Aell derivation 11}) can be used to establish (\ref{appendix, claim fast var}). Finally, if one replaces $C^{X\!X}_{\bs k}$ in (\ref{appendix, claim raw var}) with the marginal spectral density $C^{ZZobs}_{\bs k}$ then using (\ref{Aell derivation 11}) one obtains $C_{\bs\ell}^{\text{\rm var }\hat{\phi}} =  2{(2\pi)}^{-d/2} A_{\bs\ell}$.
